---
title: "21 Databases Exercises"
format: html
editor: visual
---

## 21.5.10 Exercises

1.  **What is `distinct()` translated to? How about `head()`**

    ```{r}
    flights %>%
      distinct(carrier) %>% 
      show_query()

    ```

`distinct()` function is translated to `SELECT DISTINCT`.

```{r}
flights %>% 
  head() %>% 
  show_query()
```

`head()` function is translated to `LIMIT`.

SELECT \*, distance / (air_time / 60) AS speed

FROM flights

2.  **Explain what each of the following SQL queries do and try recreate using dbplyr**

    ``` SQL
    SELECT *
    FROM flights
    WHERE dep_delay < arr_delay
    ```

    The SQL query selects all columns from the `flights` dataset where the departure delay is less than the arrival delay.

    This query can be recreatedr with:

    ```{r}
    flights %>% 
      filter(dep_delay < arr_delay) %>% 
      show_query()
    ```

    ``` SQL
    SELECT *, distance / (air_time / 60) AS speed
    FROM flights
    ```

The SQL query selects all columns from flights and adds a new column, `speed`, which calculates distance divided by air time converted converted to hours.

This can be recreated using dplyr as follows:

```{r}
flights %>% 
  mutate(speed = distance / (air_time/60)) %>% 
  show_query()
```
